### 2013-05-24 inst/examples/BUGS/process-noise-only.Rmd
Consider the case of process noise only in the parametric bayesian versions

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
library(knitcitations)
library(ggplot2) 
library(nonparametricbayes) 
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=12))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

@@

```{r jags-setup}
y <- obs$x[-1] 
N <- length(y);
jags.data <- list("N","y")
n.chains <- 3
n.iter <- 1e5
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

@@

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  theta <- exp(logtheta)
  x[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K)* (x[t] - theta) / K )
    x[t+1] ~ dnorm(mu[t], iQ) 
  }
  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "allen.bugs")
```

@@

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup")}
jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here

# Can give a list of starting values or a function that can create random values

jags.inits <- function(){
  init_p = c(r0=1, K=10, theta=2)
  list("K"=init_p["K"] * rlnorm(1,0,.1),
       "logr0"=log(rlnorm(1,0,.1) ),
       "logtheta"=log(init_p["theta"] * rlnorm(1,0,.1)), 
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
       "x"=y,        
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(1234)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- do.call(jags.parallel, list(data=jags.data, inits=jags.inits, 
                                          jags.params, n.chains=n.chains, 
                                          n.iter=n.iter, n.thin=n.thin, 
                                          n.burnin=n.burnin, model.file="allen.bugs"))

@@

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  x[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- x[t] * exp(r0 * (1 - x[t]/K) )
    x[t+1] ~ dnorm(mu[t], iQ) 
  }
  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "ricker.bugs")
```

@@

```{r ricker-mcmc, dependson="ricker-model"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=10 * rlnorm(1,0,.5),
       "logr0"=log(1) * rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags.parallel, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker.bugs"))
recompile(ricker_jags)
ricker_jags <- do.call(autojags, 
                       list(object=ricker_jags, n.update=10, n.iter=n.iter, 
                            n.thin = n.thin, progress.bar="none"))
```

@@

  x[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(x[t]), theta) / (1 + pow(abs(x[t]), theta) / K)
    x[t+1] ~ dnorm(mu[t], iQ) 
  }

  for(t in 1:(N)){
    y[t] ~ dnorm(x[t], iR)
  }
}")
writeLines(bugs.model, "myers.bugs")
```

@@

```{r myers-mcmc, dependson="myers-model"}
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list
  ("logr0"=log(1 * rlnorm(1,0,.5)), 
       "logK"=log(10* rlnorm(1,0,.5)),
       "logtheta" = log(5 * rlnorm(1,0,.5)),  
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       "x"=y,
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- do.call(jags.parallel, 
                      list(data=jags.data, inits=jags.inits, jags.params, 
                           n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                           n.burnin=n.burnin, model.file="myers.bugs")
myers_jags <- do.call(autojags, 
                      list(object=myers_jags, n.update=10, n.iter=n.iter, 
                           n.thin = n.thin, progress.bar="none"))
```

>>>>

```{r plotting-options, cache=FALSE, message=FALSE, warning=FALSE, include=FALSE}
setwd("~/Documents/code/nonparametric-bayes/inst/examples/BUGS/")
library(knitcitations)
library(ggplot2) 
library(nonparametricbayes) 
opts_chunk$set(tidy=FALSE, warning=FALSE, message=FALSE, cache=TRUE, comment=NA,
               fig.width=6, fig.height=4, cache.path="myers/", verbose=TRUE)
#opts_knit$set(upload.fun = socialR::flickr.url)
theme_set(theme_bw(base_size=12))
theme_update(panel.background = element_rect(fill = "transparent", colour = NA),
             plot.background = element_rect(fill = "transparent", colour = NA))
cbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

@@

```{r jags-setup}
y <- obs$x[-1] 
N <- length(y);
jags.data <- list("N","y")
n.chains <- 3
n.iter <- 1e4
n.burnin <- floor(10000)
n.thin <- max(1, floor(n.chains * (n.iter - n.burnin)/1000))
```

@@

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  theta <- exp(logtheta)
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K)* (y[t] - theta) / K )
    y[t+1] ~ dnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "allen_process.bugs")
```

@@

```{r allen-mcmc, dependson=c("allen-model", "allen-pars", "jags-setup"), results="hide"}
jags.params=c("K","logr0","logtheta","stdQ", "stdR") # be sensible about the order here
jags.inits <- function(){
  init_p = c(r0=1, K=10, theta=2)
  list("K"=init_p["K"] * rlnorm(1,0,.1),
       "logr0"=log(rlnorm(1,0,.1) ),
       "logtheta"=log(init_p["theta"] * rlnorm(1,0,.1)), 
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.1),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.1),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}

set.seed(1234)
# parallel refuses to take variables as arguments (e.g. n.iter = 1e5 works, but n.iter = n doesn't)
allen_jags <- do.call(jags, list(data=jags.data, inits=jags.inits, 
                                      jags.params, n.chains=n.chains, 
                                      n.iter=n.iter, n.thin=n.thin, 
                                      n.burnin=n.burnin, model.file="allen_process.bugs"))

@@

  "
  iQ <- 1 / (stdQ * stdQ);
  iR <- 1 / (stdR * stdR);
  r0 <- exp(logr0)
  y[1] ~ dunif(0, 10)
  for(t in 1:(N-1)){
    mu[t] <- y[t] * exp(r0 * (1 - y[t]/K) )
    y[t+1] ~ dnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "ricker_process.bugs")
```

@@

```{r ricker-mcmc, dependson="ricker-model"}
# Uniform priors on standard deviation terms
jags.params=c("K","logr0", "stdQ", "stdR")
jags.inits <- function(){
  list("K"=10 * rlnorm(1,0,.5),
       "logr0"=log(1) * rlnorm(1,0,.5),
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345) 
ricker_jags <- do.call(jags.parallel, 
                       list(data=jags.data, inits=jags.inits, 
                            jags.params, n.chains=n.chains, 
                            n.iter=n.iter, n.thin=n.thin, n.burnin=n.burnin,
                            model.file="ricker_process.bugs"))
recompile(ricker_jags)
ricker_jags <- do.call(autojags, 
                       list(object=ricker_jags, n.update=10, n.iter=n.iter, 
                            n.thin = n.thin, progress.bar="none"))
```

@@

  y[1] ~ dunif(0, 10)

  for(t in 1:(N-1)){
    mu[t] <- r0 * pow(abs(y[t]), theta) / (1 + pow(abs(y[t]), theta) / K)
    y[t+1] ~ dnorm(mu[t], iQ) 
  }
}")
writeLines(bugs.model, "myers_process.bugs")
```

@@

```{r myers-mcmc, dependson="myers-model"}
jags.params=c("logr0", "logtheta", "logK", "stdQ", "stdR")
jags.inits <- function(){
  list("logr0"=log(1 * rlnorm(1,0,.5)), 
       "logK"=log(10* rlnorm(1,0,.5)),
       "logtheta" = log(5 * rlnorm(1,0,.5)),  
       "stdQ"=sqrt(0.05) * rlnorm(1,0,.5),
       "stdR"=sqrt(0.1) * rlnorm(1,0,.5),
       .RNG.name="base::Wichmann-Hill", .RNG.seed=123)
}
set.seed(12345)
myers_jags <- do.call(jags.parallel, 
                      list(data=jags.data, inits=jags.inits, jags.params, 
                           n.chains=n.chains, n.iter=n.iter, n.thin=n.thin,
                           n.burnin=n.burnin, model.file="myers_process.bugs")
myers_jags <- do.call(autojags, 
                      list(myers_jags, n.update=10, n.iter=n.iter, 
                           n.thin = n.thin, progress.bar="none"))
```
